\documentclass[12pt]{article}
\usepackage{latexsym}
\usepackage{tikz}
\usepackage{tkz-graph}
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
\usetikzlibrary{shapes}

\parskip=3pt

\setlength{\textheight}{8.5in}
\setlength{\textwidth}{6in}
\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}

\newtheorem{Theorem}{Theorem}[section]
\newtheorem{Corollary}[Theorem]{Corollary}
\newtheorem{Lemma}[Theorem]{Lemma}
\newtheorem{Observation}[Theorem]{Observation}

\def\lc{\left\lceil}   
\def\rc{\right\rceil}
\def\inst#1{$^{#1}$}
\def\FAM{($\overline{K_4}$, $line$)}

\title{On k-critical ({\FAM})-free graphs}

\author{
	Dallas J. Fraser\inst{1}
	\and Ang\`ele M. Hamel'\inst{1}
	\and Ch\'inh T. Ho\`ang\inst{1}
}
\begin{document}
\maketitle

\begin{center}
{\footnotesize

\inst{1}, Department of Physics and Computer Science, Wilfrid Laurier
University, \\Waterloo, Ontario, Canada}

\end{center}

\begin{abstract}
The coloring of {\FAM}-free graph is in polynomial.

\noindent{\em Keywords}: Graph coloring, $claw$, $K_5$ - $e$, $\overline{K_4}$, $line$-$graph$ 
\end{abstract}


\section{Introduction}\label{sec:intro}
Deterimining the chromatic number of a graph is a NP-hard problem. For some graphs families the problem can be solved in polynomial time. This paper looks at $\overline{K_4}$ and $line$-free graphs. For a graph to be $line$-free is has no induced subgraph of the following graphs: $claw$, $K_5$ - $e$, $\overline{Twin-C5}$, $\overline{Twin-House}$, etc..


\section{Oberservations}\label{sec:observations}
In this section, we conclude some observations used to prove that there polytime algorithm for {\FAM}-free graphs. For the following let $G$ be a graph that is {\FAM}-free.

\begin{Lemma}\label{lem:cl-free}
$G$ contains no $C_\ell,\; \ell \geq 8$
\end{Lemma}
\noindent {\it Proof.} Since $G$ is $\overline{K_4}$-free it has no $C_\ell,\; \ell \geq 8$

\begin{Lemma}\label{lem:c7-bounded}
If $G$ contains a $C_7$ then $|G| \leq 14$
\end{Lemma}
\noindent {\it Proof.} Let $G$ contain a $C_7$. Then $G$ has no $k$-vertex for $k \in {0, 1, 2}$ since $G$ is $\overline{K_4}$-free, has no $k$-vertex for $k \in {5, 6,7}$ since $G$ is $line$-free ($claw$), and $G$ has no $k$-vertex for $k \in {3}$ since $G$ is $line$-free ($\overline{Twin-House}$). $G$ only has a $4$-vertex denoted by $Y_i$ on vertices $i,i+1,i+2,i+4$. Let $y_1$ and $y_2$ be vertices $\in Y_i,\; y_1 \neq y_2$. If $y_1y_2 \not \in E$ then there is a $claw (i+4i+5,i+4y_1,i+4y_2)$ and if $y_1y_2 \in E$ then there is a $\overline{Twin-House}(i,i+1,i+2,y_1,y_2)$ so $|Y_i| \leq 1$. Since $|Y_i| \leq 1$ $G$ can contain at most $14$ vertices.

\begin{Lemma}\label{lem:c5-kvertex}
If $G$ contains $C_5$ then $G$ has no $k$-vertex $\in {1,3,5}$
\end{Lemma}
\noindent {\it Proof.} Let $G$ contain a $C_5$. Then $G$ has no $k$-vertex for $k = 1$ since $G$ is $line$-free ($claw$), $k = 3$ since $G$ is $line$-free ($\overline{Twin-C5}$), $k = 5$ since $G$ is $line$-free ($W_5$). 

For the following let $G$ contain a $C_5$. Let the $0$-vertex set be denoted by $R$, let the $2$-vertex set on vertices $i, i+1$ be denoted by $X_i$, and let the $4$-vertex set on the vertices $i, i+1, i+2, i+3$ be denoted by $Y_i$. 

\begin{Lemma}\label{lem:c5-4vertex-bounded}
$|Y_i| \leq 1$
\end{Lemma}
\noindent {\it Proof.} Let $y_1$ and $y_2$ be vertices $\in Y_i,\; y_1 \neq y_2$. If $y_1y_2 \not \in E$ then there is a $claw (ii-1, iy_1,iy_2)$. If $y_1y_2 \in E$ then there is a $\overline{Twin-House} (i,i+1,i+2,y_1,y_2)$. So $|Y_i| \leq 5$.

\begin{Lemma}\label{lem:xi-clique}
$X_i$ forms a clique
\end{Lemma}
\noindent {\it Proof.} Let $x_1$ and $x_2$ be vertices $\in X_i,\; x_1 \neq x_2$. If $x_1x_2 \not \in E$ then there is a $claw (ii-1, ix_1, ix_2)$. So $x_1x_2 \in E$ and $X_i$ forms a clique.

\begin{Lemma}\label{lem:r-join-xi}
$R \; \circled{1} \; X_i$
\end{Lemma}
\noindent {\it Proof.} Let $x$ be a vertex $\in X_i$ and $r$ be a vertex $\in R$. If $rx \not \in E$ then there is a $\overline{K_4} (r, x, i-1, i+2)$.

\begin{Lemma}\label{lem:r-limis-xi}
If $R \neq \phi$ then $|X_i| \leq 2$
\end{Lemma}
\noindent {\it Proof.} Suppose $|X_i| \geq 3$ and $R \neq \phi$. Let $x_1,\; x_2,\;$ and $x_3$ be vertices from $X_i,\; x_1 \neq x_2,\; x_1 \neq x_3,\; x_2 \neq x_3$. By Lemma \ref{lem:xi-clique} $x_1,\; x_2,\;$ and $x_3$ forms a clique. By Lemma \ref{lem:r-join-xi} there is a $K_5$ - $e$ but $G$ is $line$-free.


\begin{center}
{\bf Acknowledgement}
\end{center}
This work was done by authors  Laurier University. The authors A.M.H. and C.T.H. were each supported by individual NSERC Discovery Grants. D.J.F was supported by an NSERC Undergraduate Student Research Award.


\clearpage
\begin{thebibliography}{99}

\end{thebibliography}

\end{document}
